Apache Flink\texttrademark: Unified Stream and Batch Processing \\ in a Single Engine

### 2015-11-26 streaming.tex
Updates from ShareLaTeX

\subsection{Stateful Stream Processing}
Simple programs in Flink’s DataStream API look like functional, side effect-free programs consisting of transformations on unbounded immutable collections. However, to cover the need for stateful operations, we incorporate explicit mutable state in the API by providing: $i)$ Operator interfaces or annotations to statically register explicit local variables within the scope of an operation;  $ii)$ an OperatorState abstraction for marking partionable dynamic state and its operations.  Flink’s checkpointing mechanism guarantees that any registered state is durably accessed exactly-once per operation via checkpointing.

As an example consider a stream wordNums containing words and an associated number for each word. We can use the partitionable  OperatorState in a stateful map function to compute a rolling average number seen per word by marking a sum and count per word as explicit states along with their updates:

\begin{lstlisting}[language=Scala]
wordNums.keyBy(0).map(new StatefulAverage())

public class StatefulAverage implements RichMapFunction<Long, Double> {
    private OperatorState<Long> sum;
    private OperatorState<Long> count;

    public Double reduce(Tuple2<String,Long> num) {
        count.update(count.value() + 1);
        sum.update(sum.value() + num.f1)
        return new Tuple2(num.f0, (double) sum.value()/count.value());
    }
...
}
\end{lstlisting}

>>>>

\subsection{Stateful Stream Processing}
While most operators in Flink’s DataStream API look like functional, side-effect free operators, they also support efficient stateful computations. Explicit mutable state is incorporated in the Flink DataStream API by providing: $i)$ Operator interfaces or annotations to statically register explicit local variables within the scope of an operation;  $ii)$ an OperatorState abstraction for marking partitioned key-value states and the associated operations. Users can also configure how the state is stored and checkpointed using the StateBackend abstractions provided by the system allowing very flexible custom state management in the streaming applications. Flink’s checkpointing mechanism guarantees that any registered state is durably accessed with exactly-once update semantics.

As a simple example consider a map operator that computes a continuous average for a given stocks' price. For this we need to keep a state with the current count and total sum for each stock in order to compute the average:

\begin{lstlisting}[language=Scala]
val priceStream: DataStream[(String,Double)] = ...
 
val avgPerStock : DataStream[(String, Double)] = source.keyBy(0).mapWithState(
    (in, state: Option[(Int, Double)]) => {
		val (count, sum) = state.getOrElse((0,0.0))
		val newState = (count + 1, sum + in._2)
		((in._1, newState._2/newState._1), newState)
	}
\end{lstlisting}
